### 2020-02-03 acknowledgements.tex
hybrid lub/ctt editted.
submit for internal rev.

  Second, I wish to extend my sincere thanks to my day-to-day advisors and mentors at KTH.
  I thank Outi Tammisolar for co-advising me since my first project, following my progress throughout the PhD
  and offering me invaluable career advice in the later years.
  I thank Jean-Christophe Loiseau for teaching me culture, programming and mathematical rigour.
  I also thank Mehdi Niazi for always being available for technical discussions or code debugging and sharing his wisdom whenever I am lost.
  The patience, support and encouragement they offered me are instrumental in building my confidence and independence, which I will never take for granted.

@@

  Finally, none of the above would have been possible without the love, hope, protection and education from my family at home.

>>>>

  Second, I wish to extend my sincere thanks to my day-to-day advisors and mentors at KTH.
  I thank Outi Tammisolar for co-advising me since my first project, following my progress throughout the PhD
  and offering me invaluable career advice in the later years.
  I thank Jean-Christophe Loiseau for instilling in me the European culture, proper programming and mathematical rigour.
  I also thank Mehdi Niazi for always being available for technical discussions or code debugging and sharing his wisdom whenever I am lost.
  The patience, support and encouragement they offered me are instrumental in building my confidence and independence, which I will never take for granted.

@@

  Finally, none of the above would have been possible without the enduring love, hope, protection and education from my family at home.

### 2020-02-03 frontmatter.tex
submit for internal rev.

The next objective of the thesis is the implementation of a minimal discrete-element lubrication/contact dynamics model for simulation of dense particle suspensions.
The two main ingredients of the model include
(i) a frame-invariant, short-range lubrication model for spherical particles, and
(ii) a soft-core, stick/slide frictional contact model activated when particles overlap.
Physically, since contact interactions dominate at high particle concentrations,
we expect the methodology to be applicable for probing the jamming of non-spherical particles and the rheology of foams as well.

@@

%
\nyckelord{droppar, suspension, flerfasflöde, mikrofluidik, mjukt material, reologi, utarmningskraft, nivåuppsättning, spökvätska, diskret element.}
%
\end{abstrakt}

>>>>

The next objective of the thesis is the implementation of a minimal discrete-element lubrication/contact dynamics model for simulation of dense particle suspensions.
The main ingredients of the model include
(i) a frame-invariant, short-range lubrication model for spherical particles, and
(ii) a soft-core, stick/slide frictional contact model activated when particles overlap.
Since contact interactions dominate at high particle concentrations,
we expect the methodology to be applicable for probing the jamming of non-spherical particles and the rheology of foams as well.

@@

\noindent
{\small (Redigerad av {\it Hanno Essén}.)}
%
\nyckelord{droppar, suspension, flerfasflöde, mikrofluidik, mjukt material, reologi, utarmningskraft, nivåuppsättning, spökvätska, diskret element.}
%
\end{abstrakt}

### 2020-02-03 overview.tex
submit for internal rev.

Droplets are also important in many artificial and engineering applications.
In material science, technicians can now routinely generate large amount of droplets in massively parallelised microfluidic channels \citep{microfluidics};
these liquid drops can serve as microcontainers, encapsulating functional substances,
and subsequently be manipulated and scaffolded into larger structures with desired properties.
In life science, droplet-based microfluidic devices have already been used to automate genome sequencing and,
in the future, may enable more effective drug development by building synthetic test platforms more faithful to the human body than
the conventional Petri dish \citep{Squires_Quakes_2005, organs-on-chips}.
In transportation, combustion engines rely on the breakup of liquid fuels into tiny drops to enhance mixing,
improve combustion efficiency, and reduce pollutant emissions.
Even in publications, printed posters, newsletters or books, including this thesis, would not have the same quality without
the precise control of pico-litre ($10^{-9}$ mL) ink droplets.

@@

This chapter summarizes the mathematical formulation and theoretical basis of microhydrodynamics and particle interactions.
By \emph{particle}, we loosely mean a solid or liquid body of micron to millimetre in size, which may also be referred to as \emph{droplet} for abuse of language.
As preliminaries, most of the contents below are given without any proof. The interested reader is encouraged to read \cite{Batchelor, hb, ps, kim_karrila, graham_2018} for greater details.

@@

Finally, we summarize the numerical algorithm of the interface-correction level set/ghost fluid method of \emph{Paper 1} in Algorithm \ref{al:ls}.
Here, the level set field is integrated in time with a total-variation-diminishing third-order Runge-Kutta scheme; the spatial derivatives, ${AD}$, are computed with either fifth-order upstream-central (HOUC) or weighted essentially non-oscillatory (WENO) schemes. The mass correction and reinitialization are treated similarly.
In the flow solver, the pressure Poisson equation is split into a constant-coefficient part and a variable-coefficient part to utilise the fast Fourier transform.
The surface tension is incorporated in the same equation using the ghost fluid method (denoted by $[\cdot]_\Gamma$).
The full algorithm is presented in details in \emph{Paper 1}. Note the high-level simplicity of the level set method comparing to the immersed boundary method (Algorithm \ref{al:ibm}).

@@

If the particle inertia is negligible, the left-hand side of Eq.\ \eqref{eq:force-balance} goes to zero,
resulting in $\bm{F}^H \approx -\bm{F}^A$ and quasi-static particle dynamics.
If the fluid inertia is also negligible, the Navier-Stokes equations reduces to the Stokes equations and
$\bm{F}^H$ may be obtained without even explicitly solving the fluid motions.
In the following, we examine two such simplified techniques.
One is more accurate and holds for any particulate system in general, albeit computationally expensive. This is the so-called \emph{Stokesian dynamics}.
The other is derived from the former, but with additional features to model particle contact; hence, it is better suited for dense suspensions.
We dub it \emph{hybrid lubrication/granular dynamics} method.

@@

Under this approximation, Eq.\ \eqref{eq:sd-multipole} can then be combined with the Fax\'{e}n's formulae for spheres (cf.\ Eq.\ \ref{eq:faxen}) to obtain a relationship between particle velocities and forces, written in matrix form as
\begin{equation} \label{eq:sd-mob-eq}
 \begin{aligned} 
  \begin{pmatrix}
   {\bm U}-{\bm U}^\infty \\
   -{\bm E}^\infty
  \end{pmatrix}
  = \mathscr{M^\infty} \cdot
  \begin{pmatrix}
   {\bm F}^H \\
   {\bm S}^H
  \end{pmatrix},
 \end{aligned}
\end{equation}
where $\mathscr{M}^\infty$ is the far-field approximation to the so-called ``grand mobility'' matrix containing the structure of all hydrodynamic interactions per particle configuration.

@@

Finally, the solution procedure of the Stokesian dynamics method can be summarized as follows.
\begin{enumerate}
\item Build the grand mobility matrix $\mathscr{M}^\infty$ from the suspension microstructure.
\item Obtain the grand resistance matrix $\mathscr{R}$ by inverting $\mathscr{M}^\infty$ and adding the lubrication correction as in Eq.\ \eqref{eq:sd-lub-correction}.
\item Compute the particle velocity according to Eq.\ \eqref{eq:sd-vel}.
\item Update the particle position using a temporal integration scheme, such as the Euler's method $\bm{x}(t+\Delta t)= \bm{x}(t) +\bm{U} \Delta t$.
\item Go back to Step 1 to continue the time integration unless the desired time step has been reached.
\end{enumerate}

@@

In a hybrid lubrication/granular dynamics method, particles interact via short-range lubrication and contact forces in a pairwise additive fashion%
\footnote{Analytically, a diverging lubrication would prevent particle contact,
though that neither feels intuitive according to everyday experience nor holds beyond the continuum limit in the presence of surface roughness.},
see Figure \ref{fig:stick-slide}.
There are several models to describe frictional contacts between particles.
The currently popular one is the stick/slide friction model employing springs and dashpots \citep{Luding2008}, which for the case of spheres reads
\begin{subequations} \label{eq:hlcd-contact}
  \begin{equation} 
    \begin{aligned}
      {\bm F}^C_{i,j} = -k_n {\bm h}_{ij}  - \gamma_n \bm{U}_{ij,n} - k_t {\bm \xi}_{ij},
    \end{aligned}
    \label{col F}
  \end{equation}
  \begin{equation} 
    \begin{aligned}
      {\bm T}^C_{i,j} = a_{i} k_t ({\bm n}_{ij} \times {\bm \xi}_{ij}),
    \end{aligned}
  \end{equation}
\end{subequations}
where ${\bm h}_{ij}=h_{ij}{\bm n}_{ij}$ denotes the signed normal surface gap between particles $i$ and $j$ (positive when overlapping),
${\bm \xi}_{ij}$ the signed tangential stretch (see \emph{Paper 7} for details, also Figure \ref{fig:stick-slide} for an illustration),
$k_n$ the normal spring constant,
$\gamma_n$ the damping constant,
and $k_t$ the tangential spring constant.
Furthermore, the contact forces need to satisfy the Coulomb's law of friction, \ie $\abs{\bm{F}^C_{tan}} \leq \mu_c\abs{\bm{F}^C_{nor}}$, with $\mu_c$ being the friction coefficient.

The above expressions for the contact forces and torques can be directly added in the ${\bm F}^A$ tuple in Eq.\ \eqref{eq:force-balance} and be solved similarly to the Stokesian dynamics method.
From a computational standpoint, the present method does not require to invert the mobility matrix and the grand resistance matrix composed of pair interactions is sparse; hence a much lower numerical cost.
Alternatively, the particle dynamics can also be obtained by explicitly solving Eq.\ \eqref{eq:force-balance}. That is, computing all forces and torques to get the acceleration, then integrating in time once to get the velocity and twice to get the position.
Similar techniques have been called dissipative particle dynamics or Brownian dynamics, depending on the context \citep{Hoogerbrugge_1992, Groot_Warren_1997}.
Provided that the particle Reynolds number remains much less than unity, this is equivalent to the quasi-static matrix inversion approach.

@@

From the fluid mechanics perspective, questions invoked by these findings naturally include,
(i) what mechanism leads to the drag reduction, and
(ii) how it can be optimised given other constraints.
The constraints involve surface chemistry, contamination, degradation, fabrication defects, wetting, etc., making the problem truly multiscale and multiphysics.
Basic research along this line, including \emph{Paper 6}, is motivated to resolve at least part of these issues.

@@

Turbulence is the unsteady flow generated from the Navier-Stokes equations at large Reynolds numbers.
As it prevails in a broad range of length and time scales, predicting the statistics of droplets or bubbles in turbulence is of great practical importance.
Take the size distribution for example.
In the ocean, air bubbles entrained in breaking waves enhance the air-sea gas flux, affect aerosol production, scavenge biological surfactants, and create ambient noise; knowing their creation mechanism and distribution in size can thus greatly improve our understanding of the nature and climate \citep{deane_stokes_2002a}.
In the atmosphere, vapor condenses on particles and forms droplets; the droplet size distribution is the defining characteristics of a cloud, controlling how it interacts with electromagnetic radiation and how fast precipitation will form \citep{Shaw_2003}.
In emulsions or colloidal suspensions, the distribution of the dispersed phase determines the stability, rheology, and area available for physical or chemical processes that occur at the interface \citep{boxall_etal_2012}. This is directly related to our daily life, as many processed food and cosmetic products are complex fluids in one way or another.

@@

Despite the ubiquity and importance, detailed numerical simulations on droplet-turbulence interactions are surprisingly scarce. It is not until the last few years that computational studies start to appear in the literature, \eg \cite{perlekar_biferale_sbragaglia_srivastava_toschi_2012a, skartlien_sollum_schumann_2013a, komrakova_eskin_derksen_2015a, scarbolo_bianco_soldati_2015a, dodd_ferrante_2016a}.
Here, the advantages of performing these large-scale simulations are dual.
On the one hand, interface-resolved, high-resolution simulations reveal a unprecedented level of details, offering arguably the full droplet and flow statistics within the afforded system. These results can then be used as benchmarks to test theoretical scalings and empirical models of the droplet distribution \citep{hinze_1955a, deane_stokes_2002a}.
On the other hand, as simulation methods for multiphase flow are continuously being developed, the current results will serve as verification for later simulations.
Based on these reasons, we performed direct numerical simulations of two immiscible and incompressible fluids in a homogeneous turbulent shear flow under various relative surface tensions (\ie Weber numbers) and initial conditions (\emph{Paper 5}).
The choice of this base flow is rationalized by its natural turbulence sustaining mechanism and the relevance to the logarithmic layer of wall-bounded turbulent flow, see \eg \cite{pumir_1996a, sekimoto_dong_jimenez_2016a}.
We admit that modelling of turbulence is difficult even by itself, let alone in the presence of droplets.
Nevertheless, numerical studies like the one we conducted shall provide additional information to future modellers and shed new lights on the droplet-turbulence interactions.

@@

The investigations presented in this thesis are motivated by the interest in understanding droplet interactions and suspension flow
that arise from a number of applications, including material science, microfluidic technology and transport of complex fluids.
The general physical picture and underlying mathematical formulation are described in Chapter \ref{chp:micro-hydro},
where we introduce elementary properties of the Stokes flow and its approximate solutions in the presence of liquid droplets or solid particles in a few simple cases.
These are the fundamentals of microhydrodynamics,
essential for developing a physical intuition and making back-of-the-envelope estimates when faced with more complicated engineering problems.
Unfortunately, sometimes a problem may be too complicated to be solved analytically, or we may simply be misguided by our intuitions.
In those cases, numerical simulations offer a relatively inexpensive means to visualise fluid motions and disentangle the various effects.
Chapter \ref{chp:num-methods} thus provides a mini-review of some currently popular methods that can be applied to simulate coupled multiphase systems.
This is the main subject of the present work --
we develop methods, implement algorithms, and simulate droplets or particles in fluids to reproduce experiments or verify theories.
The specific aim may differ by applications;
what remain the same are the numerical approach to problem solving as well as a physics-oriented search for new principles driven by curiosity.

\begin{figure}%[b]
  \centering
  \includegraphics[width=0.95\columnwidth]{paper-diagram.pdf}
  \caption{Diagram reporting the different papers in \emph{Part II}, parameterized by the number of droplets ($N$) that are or could potentially be studied,
  and by the distance (indicated by color) to being purely experimental ($-1$) or purely theoretically ($1$); 0 corresponds to purely numerical.
  The papers are labelled with short titles. Note that \emph{Paper 6} does not have a meaningful definition of $N$.}
  \label{fig:p-diagram}
\end{figure}

The detailed results of the present work are given in the next part of the thesis in the format of journal articles or technical reports.
Among them, 
papers 1--4 entail droplet interactions,
papers 5 \& 6 are miscellaneous, and
paper 7 concerns suspension flow.
Separate retrospective summaries of each paper are given in the beginning of \emph{Part II}.
A diagrammatic summary is also illustrated in Figure \ref{fig:p-diagram}.
Overall, the common limitations of the conducted studies are that simulations still fall much behind experiments in terms of physical complexity
-- which could sometimes be justified, but it is still not the same --
and the lack of simple, quantitative models that can be extracted from the results.
In general, our numerical work provides additional evidences to conclude qualitatively,
but a new set of simulations or experiments may have to be performed if precise predictions need to be made in different situations.

@@

\begin{itemize}
 \item Fabricating photonic crystals using flow-assisted droplet assembly could potentially achieve a moderate bang gap and be scaled up to meet industrial needs.
 The rest of the task is in engineering.
 Current focus of the photonic community seems to be shifted to searching for disordered hyperuniform materials
 that are easier to make and still possess bandgaps \citep{Torquato_review_2018, Joshua_2017PRL}.
 \medskip

@@

 \item Dense particle suspension is a very exciting research front,
 offers an immense parameter space to explore,
 and is related to both applied and fundamental questions.
 For example, questions like how particle geometry affects jamming,
 do active particles jam in the same way as passive particles,
 how do particle interactions and the medium topology come into play in long time limits (relevant to certain disease progression, \eg Alzheimer's disease), \etc
 are all unknown today.
 Investigating these problems requires development of new computational models and collaboration with specialists in different research areas,
 and may inspire new pathways to advance life science and biology in the future.
\end{itemize}

>>>>

Droplets are also important in many artificial and engineering applications.
In material science, technicians can now routinely generate large amounts of droplets in massively parallelised microfluidic channels \citep{microfluidics};
these liquid drops can serve as microcontainers, encapsulating functional substances,
and subsequently be manipulated and scaffolded into larger structures with desired properties.
In life science, droplet-based microfluidic devices have already been used to automate genome sequencing and,
in the future, may enable more effective drug development by building synthetic test platforms more faithful to the human body than
the conventional Petri dish \citep{Squires_Quakes_2005, organs-on-chips}.
In transportation, combustion engines rely on the breakup of liquid fuels into tiny drops to enhance mixing,
improve combustion efficiency, and reduce pollutant emissions.
Even in publications, printed posters, newsletters or books, including this thesis, would not have the same quality without
the precise control of pico-litre ($10^{-9}$ mL) ink droplets.

@@

This chapter summarises the mathematical formulation and theoretical basis of microhydrodynamics and particle interactions.
By \emph{particle}, we loosely mean a solid or liquid body of micron to millimetre in size, which may also be referred to as \emph{droplet} for abuse of language.
As preliminaries, most of the contents below are given without any proof. The interested reader is encouraged to read \cite{Batchelor, hb, ps, kim_karrila, graham_2018} for greater details.

@@

Finally, we summarise the numerical algorithm of the interface-correction level set/ghost fluid method of \emph{Paper 1} in Algorithm \ref{al:ls}.
Here, the level set field is integrated in time with a total-variation-diminishing third-order Runge-Kutta scheme; the spatial derivatives, ${AD}$, are computed with either fifth-order upstream-central (HOUC) or weighted essentially non-oscillatory (WENO) schemes. The mass correction and reinitialization are treated similarly.
In the flow solver, the pressure Poisson equation is split into a constant-coefficient part and a variable-coefficient part to utilise the fast Fourier transform.
The surface tension is incorporated in the same equation using the ghost fluid method (denoted by $[\cdot]_\Gamma$).
The full algorithm is presented in details in \emph{Paper 1}. Note the high-level simplicity of the level set method comparing to the immersed boundary method (Algorithm \ref{al:ibm}).

@@

If the particle inertia is negligible, the left-hand side of Eq.\ \eqref{eq:force-balance} goes to zero,
resulting in $\bm{F}^H \approx -\bm{F}^A$ and quasi-static particle dynamics.
If the fluid inertia is also negligible, the Navier-Stokes equations reduces to the Stokes equations and
$\bm{F}^H$ may be obtained without even explicitly solving the fluid motions.
In the following, we examine two such simplified techniques.
One is more accurate and holds for any particulate system in general, albeit computationally expensive. This is the so-called \emph{Stokesian dynamics}.
The other is derived from the former, but with additional features to model particle contact; hence, it is better suited for dense suspensions.
We dub it \emph{hybrid lubrication/granular dynamics}.

@@

Under this approximation, Eq.\ \eqref{eq:sd-multipole} can then be combined with the Fax\'{e}n's formulae for spheres (cf.\ Eq.\ \ref{eq:faxen}) to obtain a relationship between particle velocities and forces, written in matrix form as
\begin{equation} \label{eq:sd-mob-eq}
 \begin{aligned} 
  \begin{pmatrix}
   {\bm U}-{\bm U}^\infty \\
   -{\bm E}^\infty
  \end{pmatrix}
  = \mathscr{M^\infty} \cdot
  \begin{pmatrix}
   {\bm F}^H \\
   {\bm S}^H
  \end{pmatrix},
 \end{aligned}
\end{equation}
where $\mathscr{M}^\infty$ is the far-field approximation to the so-called ``grand mobility'' matrix, which contains the structure of all hydrodynamic interactions per particle configuration.

@@

Finally, the solution procedure of the Stokesian dynamics method can be summarised as follows.
\begin{enumerate}
\item Build the grand mobility matrix $\mathscr{M}^\infty$ from the suspension microstructure.
\item Obtain the grand resistance matrix $\mathscr{R}$ by inverting $\mathscr{M}^\infty$ and adding the lubrication correction as in Eq.\ \eqref{eq:sd-lub-correction}.
\item Compute the particle velocity according to Eq.\ \eqref{eq:sd-vel}.
\item Update the particle position using a temporal integration scheme, such as the Euler's method $\bm{x}(t+\Delta t)= \bm{x}(t) +\bm{U} \Delta t$.
\item Go back to Step 1 to continue the time integration unless the desired time step has been reached.
\end{enumerate}

@@

In a hybrid lubrication/granular dynamics method, particles interact via short-range lubrication and contact forces in a pairwise additive fashion%
\footnote{Analytically, a diverging lubrication would prevent particle contact,
though that neither agrees with everyday experience nor holds beyond the continuum limit in the presence of surface roughness.},
see Figure \ref{fig:stick-slide}.
There are several models to describe frictional contacts between particles.
The currently popular one is the stick/slide friction model employing springs and dashpots \citep{Luding2008}, which for the case of spheres reads
\begin{subequations} \label{eq:hlcd-contact}
  \begin{equation} 
    \begin{aligned}
      {\bm F}^C_{i,j} = -k_n {\bm h}_{ij}  - \gamma_n \bm{U}_{ij,n} - k_t {\bm \xi}_{ij},
    \end{aligned}
    \label{col F}
  \end{equation}
  \begin{equation} 
    \begin{aligned}
      {\bm T}^C_{i,j} = a_{i} k_t ({\bm n}_{ij} \times {\bm \xi}_{ij}),
    \end{aligned}
  \end{equation}
\end{subequations}
where ${\bm h}_{ij}=h_{ij}{\bm n}_{ij}$ denotes the signed normal surface gap between particles $i$ and $j$ (positive when overlapping),
${\bm \xi}_{ij}$ the signed tangential stretch (see \emph{Paper 7} for details, also Figure \ref{fig:stick-slide} for an illustration),
$k_n$ the normal spring constant,
$\gamma_n$ the damping constant,
and $k_t$ the tangential spring constant.
Furthermore, the contact forces need to satisfy the Coulomb's law of friction, \ie $\abs{\bm{F}^C_{tan}} \leq \mu_c\abs{\bm{F}^C_{nor}}$, with $\mu_c$ being the friction coefficient.

The above expressions for the contact forces and torques can be directly added in the ${\bm F}^A$ vector in Eq.\ \eqref{eq:force-balance}
and be solved similarly to the Stokesian dynamics method.
From a computational standpoint, the present method does not require to invert the mobility matrix,
and the grand resistance matrix composed of pair interactions is sparse;
hence, it can be solve much faster numerically.
Alternatively, the particle dynamics can also be obtained by explicitly solving Eq.\ \eqref{eq:force-balance}.
That is, first computing all forces and torques to get the acceleration, then integrating in time once to obtain the velocity and twice to update the position.
Similar techniques have been called dissipative particle dynamics or Brownian dynamics, depending on the context \citep{Hoogerbrugge_1992, Groot_Warren_1997}.
Provided that the particle Reynolds number remains much less than unity, this is equivalent to the quasi-static matrix inversion approach used in Stokesian dynamics.

@@

From the fluid mechanics perspective, questions invoked by these findings naturally include,
(i) what mechanism leads to the drag reduction, and
(ii) how it can be optimised given all the constraints.
The constraints involve surface chemistry, contamination, degradation, fabrication defects, wetting, etc., making the problem truly multiscale and multiphysics.
Basic research along this line, including \emph{Paper 6}, is motivated to resolve at least part of these issues.

@@

Turbulence is the unsteady flow generated from the Navier-Stokes equations at large Reynolds numbers.
As it prevails in a broad range of length and time scales, predicting the statistics of droplets or bubbles in turbulence is of great practical importance.
Take the size distribution for example.
In the ocean, air bubbles entrained in breaking waves enhance the air-sea gas flux, affect aerosol production, scavenge biological surfactants, and create ambient noise; knowing their creation mechanism and distribution in size can thus greatly improve our understanding of the nature and climate \citep{deane_stokes_2002a}.
In the atmosphere, vapor condenses on particles and forms droplets; the droplet size distribution is the defining characteristics of a cloud, controlling how it interacts with electromagnetic radiation and how fast precipitation will form \citep{Shaw_2003}.
In emulsions or colloidal suspensions, the distribution of the dispersed phase determines the stability, rheology and area available for physical or chemical processes that occur at the interface \citep{boxall_etal_2012}. This is directly related to our daily life, as many processed food and cosmetic products are complex fluids in one way or another.

@@

Despite the ubiquity and importance, detailed numerical simulations on droplet-turbulence interactions are surprisingly scarce. It is not until the last few years that computational studies start to appear in the literature, \eg \cite{perlekar_biferale_sbragaglia_srivastava_toschi_2012a, skartlien_sollum_schumann_2013a, komrakova_eskin_derksen_2015a, scarbolo_bianco_soldati_2015a, dodd_ferrante_2016a}.
Here, the advantages of performing these large-scale simulations are dual.
On the one hand, interface-resolved, high-resolution simulations reveal a unprecedented level of details, offering arguably the full droplet and flow statistics within the afforded system. These results can then be used as benchmarks to test theoretical scalings and empirical models of the droplet distribution \citep{hinze_1955a, deane_stokes_2002a}.
On the other hand, as simulation methods for multiphase flow are continuously being developed, the current results will serve as verification for later simulations.
Based on these reasons, we performed direct numerical simulations of two immiscible and incompressible fluids in a homogeneous turbulent shear flow under various relative surface tensions (\ie Weber numbers) and initial conditions (\emph{Paper 5}).
The choice of this base flow is rationalised by its natural turbulence sustaining mechanism and the relevance to the logarithmic layer of wall-bounded turbulent flow, see \eg \cite{pumir_1996a, sekimoto_dong_jimenez_2016a}.
We admit that modelling of turbulence is difficult even by itself, let alone in the presence of droplets.
Nevertheless, numerical studies like the one we conducted shall provide additional information to future modellers and shed new lights on the droplet-turbulence interactions.

@@

The investigations presented in this thesis are motivated by the interest in understanding droplet interactions and suspension flow
that arise from a number of applications, including material science, microfluidic technology and transport of complex fluids.
The general physical picture and underlying mathematical formulations are described in Chapter \ref{chp:micro-hydro},
where we introduce elementary properties of the Stokes flow and its approximate solutions in the presence of liquid droplets or solid particles in a few simple cases.
These are the fundamentals of microhydrodynamics,
essential for developing a physical intuition and making back-of-the-envelope estimates when faced with more complicated engineering problems.
Unfortunately, sometimes a problem may be too complicated to be solved by hand, or we may simply be misguided by our intuitions.
In those cases, numerical simulations offer a relatively inexpensive means to visualise fluid motions and disentangle the various effects.
Chapter \ref{chp:num-methods} thus provides a mini-review of some currently popular methods that can be applied to simulate coupled multiphase systems.
This is the main subject of the present work --
we develop methods, implement algorithms, and simulate droplets or particles in fluids to reproduce experiments or verify theories.
The specific aim may differ by applications;
what remain the same are the numerical approach to problem solving as well as a physics-oriented search for new principles driven by curiosity.

\begin{figure}%[b]
  \centering
  \includegraphics[width=0.95\columnwidth]{paper-diagram.pdf}
  \caption{Diagram reporting the different papers in \emph{Part II}, parameterized by the number of droplets ($N$) that are or could potentially be studied,
  and by the distance (indicated by color) to being purely experimental ($-1$) or purely theoretical ($1$); 0 corresponds to purely numerical.
  The papers are labelled with short titles. Note that \emph{Paper 6} does not have a meaningful definition of $N$.}
  \label{fig:p-diagram}
\end{figure}

The detailed results of the present work are given in the next part of the thesis in the format of journal articles or technical reports.
Among them, 
papers 1--4 entail droplet interactions,
papers 5 \& 6 are miscellaneous and
paper 7 concerns suspension flow.
Separate retrospective summaries of each paper are given in the beginning of \emph{Part II}.
A ``diagrammatic summary'' is also provided, as illustrated in Figure \ref{fig:p-diagram}.
Overall, the common limitations of the conducted studies are that simulations still fall much behind experiments in terms of physical complexity
-- which could be justified, sometimes --
and the lack of simple, quantitative models that can be extracted from the results.
In general, our numerical work provides additional evidences to draw qualitative conclusions,
but a new set of simulations or experiments may have to be performed if precise predictions need to be made in different situations.

@@

\begin{itemize}
 \item Fabricating photonic crystals using flow-assisted droplet assembly could potentially achieve a moderate bang gap and be scaled up to meet industrial needs.
 The rest of the task is in engineering.
 Current focus of the photonic community seems to be shifted to searching for disordered hyperuniform materials
 that are easier to make and still possess bandgaps \citep{ Joshua_2017PRL, Torquato_review_2018}.
 \medskip

@@

 \item Dense particle suspension is a very exciting research front,
 offering an immense parameter space to explore and concerning both applied and fundamental questions.
 For example, questions like how particle geometry affects jamming,
 do active particles jam in the same way as passive particles,
 how do particle interactions and the medium topology come into play in long time limits (relevant to certain disease progression, \eg Alzheimer's disease), \etc
 are all unknown today.
 Investigating these problems requires development of new computational models and collaboration with specialists in different research areas,
 which may inspire new pathways to advance life science and biology in the future.
\end{itemize}

### 2020-02-03 paper1/paper.tex
submit for internal rev.

%------------------------------------------------------------------------------
% Define title, author(s), affiliation and publishing status
%
\papertitle[ICLS/GFM for droplet suspensions under depletion forces] % Short title used in headlines (optional)
{%
  An efficient mass-preserving interface-correction level set/ghost fluid method
  for droplet suspensions under depletion forces% THE COMMENT SYMBOL AT THE END OF THIS LINE IS NEEDED
}%
%
\papertoctitle{An efficient mass-preserving interface-correction level set/ghost fluid method
  for droplet suspensions under depletion forces} % Title for toc
%
\paperauthor[Ge, Loiseau, Tammisola, Brandt] % Short authors used in headlines and List Of Papers
{%
  Zhouyang Ge, Jean-Christophe Loiseau, Outi Tammisola, Luca Brandt%
}%
%
\listpaperauthor{Z. Ge, J-Ch. Loiseau, O. Tammisola, L. Brandt}% (optional) Short authors used in List Of Papers
%
\paperaffiliation
{%
  Linn\'e FLOW Centre and SeRC, KTH Mechanics, S-100 44 Stockholm, Sweden%
}%
%
\paperjournal[J. Comput. Phys.] % Short publish info used in List Of Papers
{%
	Journal of Computational Physics%
}%
%
\papervolume{353}%
%
\papernumber{}%
%
\paperpages{435--459}%
%
\paperyear{2018}%
%
\papersummary%
{% Insert summary of the paper here (used in introduction)
In this first work, I developed and implemented an interface-correction level set/ghost fluid method (\texttt{ICLS/GFM}) for high-fidelity simulations of liquid droplets in an immiscible carrier fluid. The main novelty is a global mass correction scheme, which is conceptually easy to understand, relatively straightforward to program, and fast if implemented properly. The method works best for well-resolved droplets as validated in the paper; however, extra care must be taken when applying it otherwise. In the same paper, I also formulated a hydrodynamic model for surfactant-induced depletion forces between neighbouring drops, removing the need to simulate suspending surfactants, hence further improving the computational efficiency. The overall methodology can be readily used to study droplet interactions in liquid flows, and may be extended to simulate multiphase flows in porous media in general.
}%
%
\graphicspath{{paper1/}}%
%
%
%===============================================================================
%                            BEGIN PAPER
%===============================================================================
%
\begin{paper}

>>>>

%------------------------------------------------------------------------------
% Define title, author(s), affiliation and publishing status
%
\papertitle[ICLS/GFM for droplet suspensions under depletion forces] % Short title used in headlines (optional)
{%
  An efficient mass-preserving interface-correction level set/ghost fluid method
  for droplet suspensions under depletion forces% THE COMMENT SYMBOL AT THE END OF THIS LINE IS NEEDED
}%
%
\papertoctitle{An efficient mass-preserving interface-correction level set/ghost fluid method
  for droplet suspensions under depletion forces} % Title for toc
%
\paperauthor[Ge, Loiseau, Tammisola, Brandt] % Short authors used in headlines and List Of Papers
{%
  Zhouyang Ge, Jean-Christophe Loiseau, Outi Tammisola, Luca Brandt%
}%
%
\listpaperauthor{Z. Ge, J-Ch. Loiseau, O. Tammisola, L. Brandt}% (optional) Short authors used in List Of Papers
%
\paperaffiliation
{%
  Linn\'e FLOW Centre and SeRC, KTH Mechanics, S-100 44 Stockholm, Sweden%
}%
%
\paperjournal[J. Comput. Phys.] % Short publish info used in List Of Papers
{%
	Journal of Computational Physics%
}%
%
\papervolume{353}%
%
\papernumber{}%
%
\paperpages{435--459}%
%
\paperyear{2018}%
%
\papersummary%
{% Insert summary of the paper here (used in introduction)
In this first work, we develop, implement and validate an interface-correction level set/ghost fluid method (ICLS/GFM)
for high-fidelity simulations of liquid droplets in an immiscible carrier fluid.
The main novelty is a mass-correction scheme for global volume conservation.
It overcomes the classical ``mass loss'' issue of level set methods by constructing a correction-velocity that is easy to understand,
straightforward to program and efficient to compute numerically.
The method works best for droplets that are well-resolved as shown in the paper;
however, additional validations may be required when applying it otherwise (\eg atomisation).
In the same paper, we also formulate a hydrodynamic model for surfactant-induced depletion forces between neighbouring drops.
Our model removes the need to simulate suspending nano-scale surfactants, hence further improving the computational efficiency.
Overall, the methodology can be readily used to study droplet interactions in microfluidic devices, free-surface flows,
and may be extended to simulate multiphase flows in porous media in general.
}%
%
\graphicspath{{paper1/}}%
%
%
%===============================================================================
%                            BEGIN PAPER
%===============================================================================
%
\begin{paper}
